---
title: An R Markdown document converted from "2. CART/2. Tree-based models - CART.ipynb"
output: html_document
---

# Tree-based models - CART

In this section we will discuss the use of CART (**C**lassification **A**nd **R**egression **T**rees) and more specifically, we will use regression trees to predict claim frequencies.

```{r}
# require("CASdatasets") - Not necessary, as we saved the data as parquet file
if (!require("rpart")) install.packages("rpart")
if (!require("rpart.plot")) install.packages("rpart.plot")
if (!require("arrow")) install.packages("arrow")
if (!require("caret")) install.packages("caret")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("rattle")) install.packages("rattle")


require("rpart")
require("rpart.plot")
require("arrow")
require("caret")
require("ggplot2")
require("rattle")

options(repr.plot.width = 8, repr.plot.height = 6, repr.plot.res = 150);
```

```{r}
dataset = read_parquet(file = "../data/dataset.parquet")
```

When performing machine learning, one way to assess the predictability of the model is to use *fresh* data, not used in the *learning* process and compute some metric. This is also a good way to check whether the model is overfitting the data.

In order to have *fresh* data available at a later stage, it is common to split the dataset into two:
 - a training set. This data is used to estimate/learn the model.
 - a test set. This data is used to check the performance of the model, once it has been learned.
 
The package *caret* provides a lot of helpful functions. One of them is *createDataPartition*. Let us now split the data into two pieces (80% - 20%), by relying on stratified sampling. We use stratified sampling to obtain a similar distribution of the number of claims in boths the training and the test set. 

```{r}
set.seed(21)
in_training = createDataPartition(y = dataset$ClaimNb, times = 1, p = 0.8, list = FALSE)
training_set = dataset[in_training, ]
testing_set = dataset[-in_training, ]
```

Let us check the distribution of the variable *ClaimNb* among these two groups.

```{r}
temp = dataset
temp$training = NA
temp[in_training, "in_train_set"] = 1
temp[-in_training, "in_train_set"] = 0

summary_table = 100*round(prop.table(with(temp, table(in_train_set, ClaimNb)), 1), 5)
summary_table
rm(temp)
```

## Package CART

We will use the package CART which allows to compute regression trees. rpart can be used for regression and classification. It also implements a method for *Poisson* data.

### Quick Example

Let us start with a simple example, using two covariates:


```{r}
m0_rpart = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 method = "poisson", 
                 control = rpart.control(cp = 0.01))
summary(m0_rpart)
```

It appears that the tree has a single node and has not been split further. This comes from the complexity parameter which penalizes the splitting. **By default**, the complexity parameter **cp = 0.01**, which is often too large for Poisson data with low frequencies.

Let us put **cp = 0**, but to keep a small tree we will also impose a maximum depth of 3.

```{r}
m0_rpart = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(cp = 0, maxdepth = 3))
summary(m0_rpart)
```

The easiest way to interpret a CART is probably to plot it (if it is not too large, though!). This can be achieved with the function *rpart.plot* from the package *rpart.plot*.

```{r}
rpart.plot(m0_rpart, 
           type = 5, 
           extra = 101, 
           under = FALSE, 
           fallen.leaves = TRUE,
           digits = 3)
```

```{r}
# Or using the rattle package

fancyRpartPlot(m0_rpart, split.col="black",nn.col="black", 
               caption="", digits=2)
```

If the tree is too large, we will probably have some overfitting. To prevent overfitting, we can play with the complexity parameter cp. A good approach is to compute the whole tree, without any penalty (i.e. complexity parameter is set to 0) and afterwards prune to tree.

```{r}
m0_rpart = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(cp = 0))
rpart.plot(m0_rpart)
```

```{r}
rpart.plot(prune(m0_rpart, cp = 9e-04), 
           type = 5, 
           extra = 101, 
           under = FALSE, 
           fallen.leaves = TRUE,
           digits = 3)
```

We also see that in some terminal nodes (i.e. leaves), the number of observations (and of claims) is very low. We can set a minimum number of observation in any terminal node using minbucket

```{r}
m0_rpart = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(cp = 0, maxdepth = 3, minbucket = 1000))
rpart.plot(m0_rpart)
```

## Complexity Parameter

Playing around with the complexity parameter will yield sub-trees of the fully developped tree (i.e. the one with cp = 0).

We check on one example that increasing the complexity parameter *cp* will only prune the tree differently and that by increasing *cp*, we will obtain a **subtree**. The common parts of the three trees below (i.e., the first split) are identical.

```{r}
tree1 = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(cp = 0.0063))
rpart.plot(tree1,
           type = 5, 
           extra = 101, 
           under = FALSE, 
           fallen.leaves = TRUE,
           digits = 3)

tree2 = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(cp = 0.0013))
rpart.plot(tree2,
           type = 5, 
           extra = 101, 
           under = FALSE, 
           fallen.leaves = TRUE,
           digits = 3)


tree3 = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(cp = 0.0011))
rpart.plot(tree3,
           type = 5, 
           extra = 101, 
           under = FALSE, 
           fallen.leaves = TRUE,
           digits = 3)
```

## Cross-validation

Let us now find the optimal tree, by using cross-validation. We will again only use the variable DriverAge and CarAge in this section. By default, rpart will perform 10-fold cross-validation, using the option xval = 10. (Remark: The whole process of how the cross-validation is operated in described in Section 4.2 of rpart’s vignette: https://cran.r-project.org/web/packages/rpart/vignettes/longintro.pdf).

Essentially (and with some shortcuts), we can summarize the method as follows:

1. Estimate the full tree (cp = 0), on the whole data. We obtain a list of complexity parameters $\alpha_i$. We compute the $\beta$s are the geometric mean of two successive $\alpha_i$s. To each $\beta_i$ corresponds a subtree of the full tree.
2. Split the whole data into *folds* (generally 10 folds).
    - Estimate the tree using all folds but one. 
    - Using the list of $\beta_i$s from the previous step, we prune the tree to differents subtrees.
    - Compute the loss function ("risk") of these subtrees on the left-out fold. For each $\beta_i$ we obtain a loss.
    - Loop on the folds. We obtain for each $\beta_i$ as many losses as there are folds. We aggregate these losses by summing.
    - After agregation, for each $\beta_i$ we have a single loss.
3. The optimal complexity parameter $cp^*$ is defined as the $\beta_i$ with the smallest loss (see remark below). The optimal tree is defined as the subtree with cp = $cp^*$

Remark: We may also use the 1-SE rule, which will yield the largest $\beta_i$ whose corresponding loss is in the 1-standard error window of the smallest loss.

```{r}
m0_rpart = rpart(cbind(Exposure, ClaimNb) ~ DriverAge + CarAge, 
                 data = training_set,
                 control = rpart.control(
                             cp = 3e-5, 
                             xval = 10))
printcp(m0_rpart)
```

```{r}
plotcp(m0_rpart)
```

```{r}
head(m0_rpart$cptable, 10)
```

Let us see the optimal tree.

```{r}
cp_star = m0_rpart$cptable[which.min(m0_rpart$cptable[,"xerror"]), "CP"]
print(cp_star)

rpart.plot(prune(m0_rpart, cp = cp_star), type = 5, extra = 101, under = FALSE, fallen.leaves = FALSE,
    digits = 3)
```

Using the 1-SE error rule, we need to compute the greatest cp parameters that enters the 1-SE window

```{r}
SE = m0_rpart$cptable[which.min(m0_rpart$cptable[, 4]), 5]
min_error = m0_rpart$cptable[which.min(m0_rpart$cptable[, 4]), 4]
boundary = min_error + SE
boundary

# We need to find the first cp such that xerror < boundary.
row_optimal = which(m0_rpart$cptable[, 4] < boundary)[1]
cp_star = m0_rpart$cptable[row_optimal, 1]
print(cp_star)

rpart.plot(prune(m0_rpart, cp = cp_star), type = 5, extra = 101, under = FALSE, fallen.leaves = FALSE,
    digits = 3)
```

## Using all covariates

Let us now include all the covariates from the dataset

```{r}
m1_rpart = rpart(cbind(Exposure, ClaimNb) ~ Power + CarAge + DriverAge + Brand +
    Gas + Region + Density, data = training_set, method = "poisson", control = rpart.control(cp = 0,
    xval = 10, minbucket = 1000))
printcp(m1_rpart)
```

```{r}
plotcp(x = m1_rpart, minline = TRUE, col = "red")
```

```{r}
# Compute boundary for 1SE-rule
SE = m1_rpart$cptable[which.min(m1_rpart$cptable[, 4]), 5]
min_error = m1_rpart$cptable[which.min(m1_rpart$cptable[, 4]), 4]
boundary = min_error + SE

# Plot cross-validation error and 1SE_rule boundary.
ggplot() + geom_line(aes(x = m1_rpart$cptable[, 1], y = m1_rpart$cptable[, 4]))+
geom_point(aes(x = m1_rpart$cptable[, 1], y = m1_rpart$cptable[, 4]))+
geom_hline(aes(yintercept = boundary), color="red")+
scale_x_continuous(name = "CP") + scale_y_continuous(name = "xerror") + theme_bw()
```

If we take the value of cp that minimizes the error, we find

```{r}
cp_star = m1_rpart$cptable[which.min(m1_rpart$cptable[, 4]), 1]
cp_star
```

Let us plot the optimal tree

```{r}
m2_rpart = prune(m1_rpart, cp = cp_star)
rpart.plot(m2_rpart, type = 5, extra = 101, under = FALSE, fallen.leaves = TRUE,
    digits = 3, cex = 0.60)
```

There is a possibility to extract a variable importance metric.

```{r}
plotdata = data.frame(m2_rpart$variable.importance)
names(plotdata) = 'importance'
plotdata$var = rownames(plotdata)

ggplot(plotdata,aes(x =reorder(var,importance), y=importance)) + 
geom_bar(stat='identity')+coord_flip()+
scale_x_discrete(name="Variable") + 
theme_bw()
```

Finally, let us compute the deviance on the testing_set.

```{r}
deviance_poisson = function(x_obs, x_pred) {
    2 * (sum(dpois(x = x_obs, lambda = x_obs, log = TRUE)) - sum(dpois(x = x_obs,
        lambda = x_pred, log = TRUE)))
}

deviance_poisson(x_obs = testing_set$ClaimNb, x_pred = predict(m2_rpart, testing_set) *
    testing_set$Exposure)
```

If we compute the deviance on the full tree (not the pruned tree), we obtain

```{r}
deviance_poisson(x_obs = testing_set$ClaimNb, x_pred = predict(m1_rpart, testing_set) *
    testing_set$Exposure)
```

## Bagging of trees

Let us create the bootstrap samples.

```{r}
set.seed(85)
bootstrap_samples = createResample(training_set$ClaimNb, times = 50)
```

For each sample, we estimate a CART with the optimal complexity parameter found previously. Each tree, gives us an estimation of the claim frequency, which we average.

```{r}
bagg_cart = lapply(bootstrap_samples, function(X) {
    rpart(cbind(Exposure, ClaimNb) ~ Power + CarAge + DriverAge + Brand + Gas + Region + Density, 
          data = training_set[X, ], 
          method = "poisson", 
          control = rpart.control(cp = cp_star,
                                  xval = 0))
})
```

```{r}
pred = lapply(bagg_cart, function(X) {
    predict(X, testing_set) * testing_set$Exposure
})

pred = do.call(cbind, pred)
pred = apply(pred, 1, mean)
```

```{r}
deviance_poisson(x_obs = testing_set$ClaimNb, x_pred = pred)
```

